Generalized Tractability

### 2023-03-11 GeneralizedTractability_230310.tex
Fred M2 2023 Mac Updates

Let $\cf_d$ and $\cg_d$ Hilbert spaces, and $\SOL_d : \cf_d \to \cg_d$ be a linear solution operator with adjoint $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\[
\lambda_{\max}^2 \ge \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots,
\]
This means that the non-negative $\lambda_{i,d}$ are the singular values of the $\SOL_d$. The goal is to find an approximate solution, $\APP_d:\cb_d \times (0,\infty) \to \cg_d$ satisfying
\[
\norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon,
\]
where $\cb_d$ is the unit ball in $\cf_d$ and $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  In this case, the optimal solution is
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \SOL(u_i) \ip[\cf_d]{f}{u_i},
\]
and the complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\}.
\end{equation}

@@

\subsection{Notation and fundamental definitions}
Let $T$ be a tractability function defined as follows,
\[
T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow [1,\infty),
\]
A tractability function is a simple function that provides an upper bound on the computational complexity of a problem
A problem is $T$-tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ such that
\begin{equation} \label{eq:tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon \in (0,\infty), \ \forall d \in \mathbb{N}.
\end{equation}
A problem is \emph{strongly}
$T$-tractable with parameter $\vp$ iff the complexity is independent of the dimension, that is, there exists a positive constant $C_{\vp}$ such that
\begin{equation} \label{eq:tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall \varepsilon \in (0,\infty), \ \forall d \in \mathbb{N}.
\end{equation}

Note that the tractability function is bounded below away from zero even as $\varepsilon \to \infty$, so that the information based complexity is not expected to vanish under arbitrarily loose error requirements.  For this definition of tractability to make sense we assume that
\begin{equation}
	T \text{ is non-decreasing in all variables}
\end{equation}
which implies that the problem does not become easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit:
\begin{equation}
	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp).
\end{equation}
There is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in (1,\infty).
\end{equation}

If \eqref{eq:tract_def} holds for some $\vp$, it clearly holds for larger, $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{eq:tract_def} holds.
The  set of optimal parameters, $\mathcal{P}_{\text{opt}}$ is defined as
\begin{equation}
	\mathcal{P}_{\text{opt}} : = \{\vp^* : \eqref{eq:tract_def} \text{ holds }\forall \vp \in [\vp,\boldsymbol{\infty}) \setminus \{\vp^*\} \text{ and }
	\eqref{eq:tract_def} \text{ fails } \forall \vp \in [\vzero,\vp] \setminus \{\vp^*\} \}.
\end{equation}
In this article we allow $\vp$ to be a scalar or vector as the situation requires.

@@

\begin{theorem}\label{thm_main_strong_tract2}
A problem is strongly $T$-tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L(\vp) > 0$ such that
\begin{equation} \label{eq:strong_tractiff3}
     S_p:=\sup_{d \in \naturals} \sum_{i = L(\vp)}^\infty \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff3} holds for some $\vp$, then
\[
	\mathcal{P}_{\text{opt}} : = \{\vp^* : \eqref{eq:strong_tractiff3} \text{ holds }\forall \vp \in [\vp,\boldsymbol{\infty}) \setminus \{\vp^*\} \text{ and }
\eqref{eq:strong_tractiff3} \text{ fails } \forall \vp \in [\vzero,\vp] \setminus \{\vp^*\} \}.
\]
\end{theorem}

\begin{proof}
\textbf{Sufficient condition:}\\
Given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity:
    \begin{align*}
    \comp(\varepsilon, d) &= \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} \\
    &= \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},1,\vp)}\le \frac{1}{T(\varepsilon^{-1},1,\vp) }\biggr\}.
\end{align*}
Also since the $\lambda_{i,d}$ are non-increasing, it follows that the $T(\lambda_{i,d}^{-1},1,\vp)$ are non-decreasing. In particular, for all $n\in \mathbb{N}$, we have

\begin{align*}
    \MoveEqLeft{\lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{1,d}} \\
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},1,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},1,\vp)} \le \cdots \le \frac{1}{T(\lambda_{1, d}^{-1},1,\vp)} \\
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},1,\vp) }
    \le \frac 1n \sum_{i=1}^n  \frac{1}{T(\lambda_{i, d}^{-1},1,\vp) }
    \le \frac 1n \sum_{i=1}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}.
\end{align*}
Thus, we can conclude that
\begin{align*}
    n \ge T(\varepsilon^{-1},1,\vp) \sum_{i=1}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}
    & \implies
\frac 1n \sum_{i=1}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp) }\le  T(\varepsilon^{-1},1,\vp) \\
   & \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},1,\vp)} \le \frac{1}{T(\varepsilon^{-1},1,\vp)}.
\end{align*}

This implies an upper bound on the complexity of
\begin{align*}
       \MoveEqLeft{\comp(\varepsilon,d)} \\
       &\le 1 + T(\varepsilon^{-1},1,\vp) \sum_{i=1}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)} \\
       & = 1 + T(\varepsilon^{-1},1, \vp) \left [ \sum_{i=1}^{L(p)-1} \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}
       + \underbrace{\sum_{i=L(p)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{=:S_p <\infty\ \text{by \eqref{eq:strong_tractiff3}} }\right] \\
       & \le 1+ T(\varepsilon^{-1},1,\vp) \underbrace{\left[\frac{(L(p)-1)}{T(\lambda_{\max}^{-1},1,\vp)} + S_p\right]}_{=:C_2(p)}\\
       & = 1+ T(\varepsilon^{-1},1,\vp) C_2 (p)\\
       &\le C_{\vp} T(\varepsilon^{-1},1,\vp),
\end{align*}
for some suitably chosen $C_{\vp}>0$. This means that we have strong $T$-tractability, and verifies sufficiency of \eqref{eq:strong_tractiff3}.
As for $\mathcal{P}_{\text{opt}}$, we immediately see from the above argument that $\mathcal{P}_{\text{opt}}=(p^*,0)$, where
\[
 p^* \le \inf\{p>0\ \mbox{such that \eqref{eq:strong_tractiff3} holds}\}.
\]

\bigskip
\noindent \textbf{Necessary condition:} \\
Suppose that we have strong
$T$-tractability. That is, for some $p>0$ there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)
\]
for all $\varepsilon\in (0,\infty)$.

% This, in turn, means that
% \begin{equation}\label{eq_strong_tract_p'3}
% \comp(\varepsilon,d)\le C_{\vp} T(\varepsilon^{-1},1,p',0)
% \end{equation}
% for all $p'>p$.
%
% \noindent This means for (possibly small) $\delta_p>0$, there exists a constant $K$ such that \eqref{eq_strong_tract_p'3} holds for $p'=p+\delta_p$.

Since the eigenvalues $\lambda_{j,d}$ are non-increasing, we have
\begin{equation}\label{eq:lambda_K_strong3}
\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1,d}\le \varepsilon.
\end{equation}
Let
\[
j=j (\varepsilon):= \lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1.
\]
When we vary $\varepsilon\in (0,\infty)$, we have that $j\in \{j_d^*, j_d^*+1, j_d^*+2,\ldots\}$, where
\[
  j_d^*=\lfloor  T(0,1,\vp)\rfloor +1,
\]
by the monotonicity of $T$.

Note furthermore that we always have
\[
j=j(\varepsilon)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1.
\]
Due to our assumptions, we know that
$T(\cdot,1, \vp)$ is increasing in the first argument if
we keep $p$ fixed.

Then, by \eqref{eq:lambda_K_strong3},
\[
j \le C_{\vp} T(\varepsilon^{-1},1,\vp)+1\le C_{\vp} T(\lambda_{j,d}^{-1},1,\vp)+1,
\]
and the latter chain of inequalities holds for all $j\ge j_d^*$.

This implies
\[
 T(\lambda_{j,d}^{-1},1, \vp) \ge \frac{j-1}{C_{\vp}}.
\]
This, in turn, implies, for any $\tau>1$, that
\[
 (T(\lambda_{j,d}^{-1},1, \vp))^\tau \ge \frac{(j-1)^\tau}{C_{\vp}^\tau },
\]
which yields, due to our assumption on $T$,
\[
 K_{\vp,\tau}\,T (\lambda_{j,d}^{-1},1,\tau \vp) \ge \frac{(j-1)^\tau}{C_{\vp}^\tau},
\]
for some $K_{\vp,\tau}>0$.
This gives
\[
\sum_{j=j_d^*}^\infty \frac{1}{T (\lambda_{j,d}^{-1},1, \tau \vp)} \le K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{j=j_d^*}^\infty \frac{1}{(j-1)^\tau} \le
\peter{K_{\vp,\tau}\, C_{\vp}^\tau\, \zeta (\tau) }< \infty,
\]
\peter{where by $\zeta(\cdot)$ we denote the Riemann zeta function. Consequently,
\[
\sup_{d\in\naturals} \sum_{j=j_d^*}^\infty \frac{1}{T (\lambda_{j,d}^{-1},1, \tau \vp)} \le K_{\vp,\tau}\, C_{\vp}^\tau\, \zeta (\tau) < \infty.
\]
}

Note, however, that, \peter{for any $d\in\naturals$,}
\[
 j_d^* = \lfloor C_{\vp}\, T(0,1,\vp)\rfloor +1 \le \lfloor C_{\vp}\, T(0,1,\tau \vp)\rfloor +1
\]
Hence, we can find $L(\tau p)$, which is independent of $\varepsilon$ and $d$, such that
\[
 \peter{\sup_{d\in\naturals}} \sum_{j=L(\tau p)}^\infty \frac{1}{T(\lambda_{j,d}^{-1},1, \tau \vp)}<\infty.
\]
This yields \eqref{eq:strong_tractiff3} with $p$ replaced by $\tau p$, and so we see necessity of $\eqref{eq:strong_tractiff3}$.

@@

\end{proof}
\bigskip

>>>>

Let $\cf_d$ and $\cg_d$ Hilbert spaces, and $\SOL_d : \cf_d \to \cg_d$ be a linear solution operator with adjoint $\SOL_d^*$ such that $\SOL_d^*\SOL_d : \cf_d \to \cf_d$ has eigenvalues and orthonormal eigenvectors
\[
\lambda_{\max}^2 \ge \lambda_{1,d}^2 \ge \lambda_{2,d}^2 \ge \cdots, \qquad u_{1,d}, u_{2,d}, \ldots,
\]
This means that the non-negative $\lambda_{i,d}$ are the singular values of the $\SOL_d$. To avoid trivial cases, we assume that there are an infinite number of positive eigenvalues.

The goal is to find an approximate solution, $\APP_d:\cb_d \times (0,\infty) \to \cg_d$ satisfying
\[
\norm[\cg_d]{\SOL_d(f) - \APP_d(f,\varepsilon)} \le \varepsilon,
\]
where $\cb_d$ is the unit ball in $\cf_d$ and $\APP_d(f,\varepsilon)$ is allowed to depend on arbitrary linear functionals.  In this case, the optimal solution is
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \SOL(u_i) \ip[\cf_d]{f}{u_i},
\]
and the complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\}.
\end{equation}

@@

\subsection{Notation and fundamental definitions}
Let $T$ be a tractability function defined as follows,
\[
T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow [1,\infty),
\]
A tractability function is a simple function that provides an upper bound on the computational complexity of a problem
A problem is $T$-tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ such that
\begin{equation} \label{eq:tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},d,\vp) \qquad \forall \varepsilon >0, \ d \in \mathbb{N}.
\end{equation}
A problem is \emph{strongly}
$T$-tractable with parameter $\vp$ iff the complexity is independent of the dimension, that is, there exists a positive constant $C_{\vp}$ such that
\begin{equation} \label{eq:tract_def}
	\comp(\varepsilon,d) \le C_{\vp}\, T(\varepsilon^{-1},1,\vp) \qquad \forall \varepsilon >0 , \ d \in \mathbb{N}.
\end{equation}

Note that the tractability function is bounded below away from zero even as $\varepsilon \to \infty$, so that the information based complexity is not expected to vanish under arbitrarily loose error requirements.  For this definition of tractability to make sense we assume that
\begin{subequations} \label{eq:Tconditions}
\begin{equation}
	T \text{ is non-decreasing in all variables}
\end{equation}
which implies that the problem does not become easier by decreasing the tolerance, $\varepsilon$, or increasing the dimension, $d$. Since there are an infinite number of positive eigenvalues, it also make sense to assume that
\begin{equation}
	\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d,\vp) = \infty \qquad \forall d \in \naturals, \ \vp \in [0,\infty)^s.
\end{equation}
Since $T(\cdot,d,\vp)$ is non-decreasing, we may define the following limit:
\begin{equation}
	T(0,d,\vp):=\lim_{\varepsilon\to\infty}T(\varepsilon^{-1},d,\vp).
\end{equation}
There is a technical assumption required for our analysis.  There exists a $K_{\vp,\tau}$ depending on $\vp$ and $\tau$, but  independent of $\varepsilon$ and $d$, such that
\begin{equation} \label{eq:ptauassume}
	[T(\varepsilon^{-1},d,\vp)]^\tau \le K_{\vp,\tau} T(\varepsilon^{-1},d,\tau \vp),   \quad \forall \varepsilon \in (0,\infty), \ d \in \naturals, \ \vp\in[0,\infty)^s, \ \tau\in (1,\infty).
\end{equation}
\end{subequations}

If \eqref{eq:tract_def} holds for some $\vp$, it clearly holds for larger, $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{eq:tract_def} holds.
The  set of optimal parameters, $\mathcal{P}_{\text{opt}}$ is defined as
\begin{equation}
	\mathcal{P}_{\textup{opt}} : = \{\vp^* : \eqref{eq:tract_def} \text{ holds }\forall \vp \in [\vp,\boldsymbol{\infty}) \setminus \{\vp^*\} \text{ and }
	\eqref{eq:tract_def} \text{ fails } \forall \vp \in [\vzero,\vp] \setminus \{\vp^*\} \}.
\end{equation}
In this article we allow $\vp$ to be a scalar or vector as the situation requires.

@@

\begin{theorem}\label{thm_main_strong_tract2}
A problem is strongly $T$-tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L(\vp) > 0$ such that
\begin{equation} \label{eq:strong_tractiff3}
     S(p):=\sup_{d \in \naturals} \sum_{i = L(\vp)}^\infty \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff3} holds for some $\vp$, then
\[
	\widetilde{\mathcal{P}}_{\textup{opt}} := \{\vp^* : \eqref{eq:strong_tractiff3} \text{ holds }\forall \vp \in [\vp,\boldsymbol{\infty}) \setminus \{\vp^*\} \text{ and }
\eqref{eq:strong_tractiff3} \text{ fails } \forall \vp \in [\vzero,\vp] \setminus \{\vp^*\} \} = \mathcal{P}}_{\textup{opt}.
\]
\end{theorem}

\begin{proof}
\textbf{Sufficient condition:}\\
Given that $T$ is increasing in its arguments, we have the following equivalent expression for the complexity:
    \begin{equation*}
    \comp(\varepsilon, d) &= \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\} = \min \biggl\{n \in \natzero : \frac{1}{T(\lambda_{n+1,d}^{-1},1,\vp)}\le \frac{1}{T(\varepsilon^{-1},1,\vp) }\biggr\}.
\end{equation*}
Also since the $\lambda_{i,d}$ are non-increasing, it follows that the $T(\lambda_{i,d}^{-1},1,\vp)$ are non-decreasing. In particular, for all $n\in \mathbb{N}$, we have

\begin{align*}
    \MoveEqLeft{\lambda_{n+1,d} \le \lambda_{n,d} \le \cdots \le \lambda_{1,d}} \\
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},1,\vp)} \le \frac{1}{T(\lambda_{n, d}^{-1},1,\vp)} \le \cdots \le \frac{1}{T(\lambda_{1, d}^{-1},1,\vp)} \\
    & \implies \frac{1}{T(\lambda_{n+1, d}^{-1},1,\vp) }
    \le \frac 1n \sum_{i=1}^n  \frac{1}{T(\lambda_{i, d}^{-1},1,\vp) }
    \le \frac 1n \sum_{i=1}^\infty  \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}.
\end{align*}
Thus, we can conclude that
\begin{equation*}
    n \ge T(\varepsilon^{-1},1,\vp) \sum_{i=1}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}
   \implies   \frac{1}{T(\lambda_{n+1, d}^{-1},1,\vp)} \le
   \frac 1n \sum_{i=1}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp) } \le \frac{1}{T(\varepsilon^{-1},1,\vp)}.
\end{equation*}

This inequality implies an upper bound on the complexity of
\begin{align*}
       \comp(\varepsilon,d)
       &\le 1 + T(\varepsilon^{-1},1,\vp) \sum_{i=1}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)} \\
       & = 1 + T(\varepsilon^{-1},1, \vp) \left [ \sum_{i=1}^{L(p)-1} \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}
       + \underbrace{\sum_{i=L(p)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},1,\vp)}}_{\le S(p) <\infty\ \text{by \eqref{eq:strong_tractiff3}} }\right] \\
       & \le 1+ T(\varepsilon^{-1},1,\vp) \underbrace{\left[\frac{(L(p)-1)}{T(\lambda_{\max}^{-1},1,\vp)} + S(p)\right]}_{=:C_1(p)}\\
       & = 1+ T(\varepsilon^{-1},1,\vp) C_1 (p)\\
       &\le C_{\vp} T(\varepsilon^{-1},1,\vp),
\end{align*}
for some suitably chosen $C_{\vp}>0$. This means that we have strong $T$-tractability, and verifies sufficiency of \eqref{eq:strong_tractiff3}.

\bigskip
\noindent \textbf{Necessary condition:} \\
Suppose that we have strong
$T$-tractability. That is, for some $\vp \ge \vzero$ there exists a positive constant $C_{\vp}$ such that
\[
\comp(\varepsilon,d)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)
\qquad \forall \varepsilon > 0, \ d \in \mathbb{N}.
\]
Since the eigenvalues $\lambda_{j,d}$ are non-increasing, we have
\begin{equation}\label{eq:lambda_K_strong3}
\lambda_{\lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1,d}\le \varepsilon.
\end{equation}

Define
\[
j (\varepsilon,\vp):= \lfloor C_{\vp}\, T(\varepsilon^{-1},1,\vp)\rfloor +1, \quad
 L(\vp):= j (\infty,\vp).
\]
Thus, it follows by \eqref{eq:lambda_K_strong3} that $\lambda_{j(\varepsilon,\vp),d} \le \varepsilon$.
Note furthermore that we always have
\[
j(\varepsilon)\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1 \le C_{\vp} T(\lambda_{j(\varepsilon,\vp),d}^{-1},1,\vp)+1, \qquad \forall \varepsilon > 0,
\]
since
$T(\cdot,1, \vp)$ is non-decreasing.

For $\varepsilon$ taking on all positive values, $j(\varepsilon,\vp)$ takes on all values greater than or equal to $L(\vp)$, so
\[
j\le C_{\vp}\, T(\varepsilon^{-1},1,\vp)+1 \le C_{\vp} T(\lambda_{j,d}^{-1},1,\vp)+1, \qquad \forall j \ge L(\vp).
\]
This implies via our technical assumption \eqref{eq:ptauassume} that
\begin{equation*}
 K_{\vp,\tau}\,T (\lambda_{j,d}^{-1},1,\tau \vp) \ge
 [T(\lambda_{j,d}^{-1},1, \vp)]^\tau
 \ge
  \left[\frac{(j-1)}{C_{\vp}}\right]^\tau \qquad \forall j \ge L(\vp)
\end{equation*}
Summing both sides of the inequality yields and noting that $L(\vp) \le L(\tau\v)$, it follows that
\begin{multline*}
\sup_{d\in\naturals} \sum_{j=L(\tau\vp)}^\infty \frac{1}{T (\lambda_{j,d}^{-1},1, \tau \vp)} \le
\sup_{d\in\naturals} \sum_{j=L(\vp)}^\infty \frac{1}{T (\lambda_{j,d}^{-1},1, \tau \vp)} \\
\le K_{\vp,\tau}\, C_{\vp}^\tau
\sum_{j=L(\vp)}^\infty \frac{1}{(j-1)^\tau} \le
K_{\vp,\tau}\, C_{\vp}^\tau\, \zeta (\tau) < \infty,
\end{multline*}
where $\zeta$ denotes the Riemann zeta function.
This yields \eqref{eq:strong_tractiff3} with $p$ replaced by $\tau p$, and so we see the necessity of $\eqref{eq:strong_tractiff3}$.

\bigskip
\noindent \textbf{Optimality:} \\
To complete the proof, we must show that $\widetilde{\cp}_{\textup{opt}} = \cp_{\textup{opt}}$.

@@

As for $\mathcal{P}_{\text{opt}}$, we immediately see from the above argument that $\mathcal{P}_{\text{opt}}=(p^*,0)$, where
\[
p^* \le \inf\{p>0\ \mbox{such that \eqref{eq:strong_tractiff3} holds}\}.
\]
